---
title: "1. Statsbomb Data"
author: "Ekrem Bayar"
date: "25 05 2020"
output: html_document
---

# 1. Packages

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(rjson)
library(parallel)
library(DT)
```

# 2. Data

# 2.1. Competitions

competitions.json files includes tournament details.

- La Liga 2004-2019 Seasons
- England FA Women's Super League 2018-2020 Seasons
- Women's World Cup 2019
- FIFA World Cup 2018
- USA Women's League 2018 (NWSL)

```{r}
competitions <- fromJSON(file="C:\\Users\\dell\\Desktop\\FoT\\open-data-master\\data\\competitions.json")

competitions <- bind_rows(competitions)

datatable(competitions)
```

## 2.2. Matches

matches file gives us match details with json format.

```{r}
match.files <- list.files(path="C:\\Users\\dell\\Desktop\\FoT\\open-data-master\\data\\matches",
                          full.names = TRUE,recursive = TRUE)
match.files
```

```{r}
all.matches <- lapply(match.files, function(x){
  
  match.temp <- fromJSON(file=x)
  
  matches.df <- lapply(match.temp, function(y){
    
    data.frame(t(unlist(y)), stringsAsFactors = FALSE)
    
  })
  
  bind_rows(matches.df)
  
}) %>% bind_rows()

rm(match.files)

datatable(all.matches)
```

### 2.2.1. Number of Matches in each league by season   
```{r}
all.matches %>% 
  group_by(competition.competition_name, season.season_name, home_team.home_team_gender) %>% 
  count() %>% 
  datatable()
```

- There are 64 matches in the FIFA World Cup.
- FA Women's Super League	2018/2019 includes 107 matches but In the season had played 110 matches.
- FA Women's Super League	2019/2020, Wikipedia shows us 82 matches but there are 87 matches here.
- Women's World Cup	2019 is also usable. It has 52 matches. 

### 2.2.2. Number of Matches in each team by season 
```{r}
numofmatches <- rbind(
  all.matches %>% select(competition.competition_name, season.season_name, 
                         home_team.home_team_name, home_team.home_team_gender) %>% 
    rename(competition = competition.competition_name, gender = home_team.home_team_gender,
           team = home_team.home_team_name, season = season.season_name),
  
  all.matches %>% select(competition.competition_name, season.season_name, 
                         away_team.away_team_name, home_team.home_team_gender) %>% 
    rename(competition = competition.competition_name, gender = home_team.home_team_gender,
           team = away_team.away_team_name, season = season.season_name)
) %>% group_by(competition, season, team, gender) %>% 
  count()

datatable(numofmatches)
```


There are 107 matches in the FA Women's Super League 2018/2019 but it had to be 110 matches. There are no 3 matches details in first six observation.


```{r}
numofmatches %>% 
  filter(competition == "FA Women's Super League", season == "2018/2019") %>% 
  arrange(n) %>% 
  datatable()
```

In La Liga, Barcelona has some missings.
```{r}
numofmatches %>% 
  filter(team == "Barcelona") %>% 
  datatable()
```

```{r}
numofmatches %>% 
  filter(competition == "La Liga") %>% arrange(-n) %>% 
  datatable()
```

```{r}
numofmatches %>% 
  filter(competition == "FIFA World Cup") %>% arrange(-n) %>% 
  datatable()
```

```{r}
all.matches %>% 
  group_by(competition.competition_name, season.season_name, 
           away_team.away_team_name, home_team.home_team_gender) %>% 
  count() %>% 
  datatable()
```


# 2.3. Lineups

Lineups gives player details, starting eleven and bench in each team.

```{r}
lineups <- list.files(path="C:\\Users\\dell\\Desktop\\FoT\\open-data-master\\data\\lineups",
                          full.names = TRUE,recursive = TRUE)
head(lineups)
```

```{r}
lineups <- lapply(lineups, function(x){
  # Match Id
  match_id <- data.frame(match_id = as.numeric(str_extract_all(x, "\\(?[0-9]+\\)?")[[1]]))
  # Read Json
  temp <- fromJSON(file=x)
  
  
  team <- lapply(temp, function(y){
    
    lineups_list <- y$lineup
      
    # Match Id & Name
    team_details <- data.frame(
      team_id = y$team_id,
      team_name = y$team_name,
      stringsAsFactors = FALSE
      )
    
    lineups_list <- lapply(lineups_list, function(z){
      
      if(is.null(z$player_nickname)){nickname <- NA_character_}else{nickname <- z$player_nickname}
      if(is.null(z$jersey_number)){jersey_number <- NA_integer_}else{jersey_number <- z$jersey_number}
      if(is.null(z$country$id)){country_id <- NA_integer_}else{country_id <- z$country$id}
      if(is.null(z$country$name)){country_name <- NA_character_}else{country_name <- z$country$name}
      
      data.frame(
        player_id = z$player_id,
        player_name = z$player_name,
        player_nickname = nickname,
        jersey_number = jersey_number,
        country_id = country_id,
        country_name = country_name,
        stringsAsFactors = FALSE
      )
      
    }) %>% bind_rows()
    
    cbind(team_details, lineups_list)
    
  }) %>% bind_rows()
  
  cbind(match_id, team)
  
}) %>% bind_rows() %>% mutate(player_nickname = if_else(is.na(player_nickname), player_name, player_nickname))

datatable(lineups[1:100,])
```



# 2.4. Maç Özeti: Events 

Event Data gives ball actions in the match.

```{r}
event.files <- list.files(path="C:\\Users\\dell\\Desktop\\FoT\\open-data-master\\data\\events",
                          full.names = TRUE,recursive = TRUE)
head(event.files)
```

### 2.4.1. FA Women's Super League Matches

FA Women's Super League has more suitable data

```{r}
wmi <- all.matches %>% 
  filter(competition.competition_name == "FA Women's Super League",	season.season_name == "2018/2019") %>% 
  pull(match_id)

head(wmi)

wmi_index <- data.frame(
  index = 1:length(event.files),
  event.files = event.files,
  stringsAsFactors = FALSE
) %>% filter(unlist(str_extract_all(event.files, "\\(?[0-9]+\\)?")) %in% wmi) %>% 
  pull(index)

head(event.files[wmi_index])
```



```{r}
events <- mclapply(event.files[wmi_index], function(x){
  
  match_id <- data.frame(match_id = as.numeric(str_extract_all(x, "\\(?[0-9]+\\)?")[[1]]))
  
  temp <- fromJSON(file=x)
  
  event <- mclapply(temp, function(y){
    
    data.frame(
      id = y$id,
      index = y$index,
      event = y$type$name,
      play.pattern = y$play_pattern$name,
      period = y$period,
      timestamp = y$timestamp,
      minute = y$minute,
      second = y$second,
      possession = y$possession,
      possession.team = y$possession_team$name,
      team = y$team$name,
      stringsAsFactors = FALSE
    )
  }) %>% bind_rows()
  
  cbind(match_id, event)

}) %>% bind_rows()
```









































