---
title: "6. Statsbomb Shot & Pass Maps"
author: "Ekrem Bayar"
date: "25 05 2020"
output: html_document
---

# 1. Friends of Tracking: Making Your Own Shot and Pass Maps

*Video:* https://www.youtube.com/watch?v=oOAnERLiN5U

*Github:* https://github.com/Friends-of-Tracking-Data-FoTD/SoccermaticsForPython/blob/master/2PlotShotsAndPasses.py


```{r}
user = "erhan"


if(user == "erhan")
  {path = "D:\\Akademik ドlmalar\\Spor Bilimi\\Futbol\\Statsbomb\\data\\"}
if(user == "ekrem")
  {path = "C:\\Users\\dell\\Desktop\\FoT\\open-data-master\\data\\"}

```


# 2. Packages
```{r message=FALSE, warning=FALSE}
library(plyr)
library(tidyverse)
library(magrittr)
library(parallel)
library(rjson)
```

# 3. Pitch 
```{r}
source(file = "Functions/Pitch for Statsbomb.R")
```


# 3. Shots: Women's World Cup 19: England vs Sweden

```{r}
#ID for England vs Sweden Womens World Cup
# match id = 69301

df <- fromJSON(file=paste0(path,"\\events\\69301.json"))

# Shot Index
shot.index <- which(sapply(df,function(x) x$type$name)=="Shot")
shot.index


# Json Format覺n覺 Direk Data Frame'e D繹n羹t羹rme
shots <- mclapply(df[shot.index], function(x){ unlist(x)})

library(plyr)
shots <- rbind.fill(mclapply(shots, 
    function(x) do.call("data.frame", as.list(x))
))



shots %<>% select(-contains("freeze")) %>% 
  mutate_if(is.factor, as.character) %>% 
  mutate_at(vars(c(location1, location2, shot.statsbomb_xg)), funs(as.numeric))
```



## Visualization

```{r}
home_team ="England Women's"
away_team ="Sweden Women's"

sut <- shots %>%  
  mutate(location2 = if_else(team.name == home_team,  80 - location2, location2),
         location1 = if_else(team.name == away_team, 120 - location1, location1)) 


# Shots
pitch()+
  geom_point(sut %>% filter(shot.outcome.name == "Goal"), 
             mapping = aes(location1, location2, color = team.name), size = 5)+
  geom_point(sut %>% filter(shot.outcome.name != "Goal"),
             mapping = aes(location1, location2, color = team.name), size = 5, alpha = 0.4)+
  scale_color_manual(values = c("red", "blue"))+
  geom_text(sut %>% filter(shot.outcome.name == "Goal"), 
            mapping = aes(location1, location2, label = player.name),vjust = -0.5, hjust = -0.01)+
  labs(title = "England Women's Shots  |  Sweden Women's Shots", color = NULL)+
  theme(
    legend.position = "bottom",
    plot.title = element_text(hjust=0.5, vjust = -5)
    )




##### XG'羹n neden karesini al覺yor oray覺 tam anlamad覺m eitimde

# Expected Goal
pitch()+
  geom_point(sut %>% filter(shot.outcome.name == "Goal"), 
             mapping = aes(location1, location2, color = team.name,
                           size = sqrt(shot.statsbomb_xg*15)))+
  geom_point(sut %>% filter(shot.outcome.name != "Goal"), 
             mapping = aes(location1, location2, color = team.name,  
                           size = sqrt(shot.statsbomb_xg*15)),  alpha = 0.4)+
  scale_color_manual(values = c("red", "blue"))+
  geom_text(sut %>% filter(shot.outcome.name == "Goal"), 
            mapping = aes(location1, location2, label = player.name),
            vjust = -0.5, hjust = -0.01)+
  labs(title = "England Women's Shots  |  Sweden Women's Shots", color = NULL, size = "XG")+
  theme(
    legend.position = "bottom",
    plot.title = element_text(hjust=0.5, vjust = -5)
    )

```


# 4. Passes: Women's World Cup 19: England vs Sweden

> NOT: Sara Caroline Seger'in pas覺 yanl覺 癟覺k覺yor bende videoda farkl覺!


## 4.1. Create a dataframe of passes which contains all the passes in the match

```{r}
# Pass Index
pass.index <- which(sapply(df,function(x) x$type$name)=="Pass")
pass.index %>% head


# Json Format覺n覺 Direk Data Frame'e D繹n羹t羹rme
pass <- mclapply(df[pass.index], function(x){ unlist(x)})

library(plyr)
pass <- rbind.fill(mclapply(pass, 
    function(x) do.call("data.frame", as.list(x))
))

head(pass)
```


## 4.2 Plot the start point of every Sweden pass. Attacking left to right.


```{r}
sweden <- pass %>% 
  filter(team.name == "Sweden Women's") %>% 
  mutate_if(is.factor, as.character) %>% 
  mutate_at(vars(c(location1, location2)), funs(as.numeric)) 
  

pitch()+
  geom_point(sweden, mapping = aes(120-location1, location2))+
  labs(title = "Sweden Women's All Pass Locations")
```



## 4.3. Plot only passes made by Caroline Seger (she is Sara Caroline Seger in the database)

pitchLengthX=120
pitchWidthY=80

```{r}
sara.caroline.seger <- pass %>% 
  filter(player.name == "Sara Caroline Seger") %>% 
  mutate_if(is.factor, as.character) %>% 
  mutate_at(vars(c(location1, location2, pass.end_location1, pass.end_location2)), funs(as.numeric)) %>% 
  mutate(dx = pass.end_location1 - location1, 
         dy = pass.end_location2 - location2)

pitch()+
  geom_point(sara.caroline.seger, mapping = aes(location1, 80-location2))+
  labs(title = "Sara Caroline Seger's All Pass Locations",
       subtitle = paste("S. Caroline Seger:", unique(sara.caroline.seger$position.name)))



pitch()+
  geom_point(sara.caroline.seger, mapping = aes(location1, 80-location2), size = 4, color = "blue")+
  geom_point(sara.caroline.seger, mapping = aes(location1, 80-location2, label = index), size = 4)+
  geom_segment(sara.caroline.seger, mapping = aes(x = location1, y = 80-location2,
                                                  xend = pass.end_location1, yend = 80-pass.end_location2),
               color = "yellow", arrow = arrow(length = unit(0.01, "npc")))


```



#4, Plot arrows to show where the passes we

```{r}
pitch()+
  geom_point(sara.caroline.seger, mapping = aes(location1, location2), color = "blue")+
  geom_segment(sara.caroline.seger, mapping = aes(x = location1, xend = pass.end_location1, y = location2, yend = pass.end_location2),
               arrow = arrow(length = unit(0.03, "npc")), color = "yellow")+
  labs(title = "Sara Caroline Seger's All Pass Locations",
       subtitle = paste("S. Caroline Seger:", unique(sara.caroline.seger$position.name)))
```


```{r}
pitch()+
  geom_point(sara.caroline.seger, mapping = aes(location1, 80-location2), color = "blue")+
  geom_segment(sara.caroline.seger, 
               mapping = aes(x = location1, xend = pass.end_location1-location1, 
                             y = location2, yend = pass.end_location2-location2))
```


# Homework: France - Argentina


```{r}
#ID for England vs Sweden Womens World Cup
# match id = 7551

fr_ar <- fromJSON(file=paste0(path,"events\\7580.json"))

sapply(fr_ar,function(x) x$player$name) %>% unlist() %>% unique()


# Shot Index
player.index <- which(sapply(fr_ar,function(x) x$player$name)=="Lionel Andr矇s Messi Cuccittini")

# Json Format覺n覺 Direk Data Frame'e D繹n羹t羹rme
messi <- lapply(fr_ar[player.index], function(x){ unlist(x)})
library(plyr)
messi <- rbind.fill(lapply(messi, 
    function(x) do.call("data.frame", as.list(x))
))


messi %<>% 
  mutate_if(is.factor, as.character) %>% 
  mutate_at(vars(c(location1, location2, pass.end_location1, pass.end_location2)), funs(as.numeric)) 


messi %>% dplyr::count(type.name) %>% arrange(-n)
```



```{r}
messi %>% filter(type.name == "Pass") %>% select(contains("pass"))
```


```{r}
pitch(pitch_theme = "blue")+
  geom_point(messi %>% filter(type.name == "Pass"), mapping = aes(location1, location2, color = pass.height.name, shape = play_pattern.name), size = 5)+
  geom_segment(messi, mapping = aes(x = location1, y = location2, xend = pass.end_location1, yend = pass.end_location2), 
               arrow = arrow(length = unit(0.03, "npc")))+
  scale_color_manual(values = c("green","yellow","red"))


# Coordflip eklendiinde doru oluyor gibi
pitch(pitch_theme = "blue")+
  geom_point(messi %>% filter(type.name == "Pass"), mapping = aes(location1, location2, color = pass.height.name, shape = play_pattern.name), size = 5)+
  geom_segment(messi, mapping = aes(x = location1, y = location2, xend = pass.end_location1, yend = pass.end_location2), 
               arrow = arrow(length = unit(0.03, "npc")))+
  scale_color_manual(values = c("green","yellow","red"))+
  coord_flip()

# Aa覺daki de doru dikeyde coordfilip

pitch(pitch_theme = "blue")+
  geom_point(messi %>% filter(type.name == "Pass"), mapping = aes(location1, 80-location2, color = pass.height.name, shape = play_pattern.name), size = 5)+
  geom_segment(messi, mapping = aes(x = location1, y = 80-location2, xend = pass.end_location1, yend = 80-pass.end_location2, color = pass.height.name), 
               arrow = arrow(length = unit(0.03, "npc")))+
  scale_color_manual(values = c("green","yellow","red"))+
  scale_shape_manual(values = c(17, 15,3,4,20))+
  labs(title = "Messi's Pass Map", subtitle = "FIFA World Cup 18: France-Argentina", shape = "Play Pattern", color = "Height")+
  theme(text = element_text(color = "white"),
        plot.title = element_text(hjust=0.5, vjust = -5),
        plot.subtitle = element_text(hjust=0.5, vjust = -5))







pitch(pitch_theme = "blue")+
  geom_point(messi %>% filter(type.name == "Pass"), mapping = aes(location1, location2, color = pass.height.name, shape = play_pattern.name), size = 5)+
  geom_segment(messi, mapping = aes(x = location1, y = location2, xend = pass.end_location1, yend = pass.end_location2,color = pass.height.name), 
               arrow = arrow(length = unit(0.03, "npc")))+
  scale_color_manual(values = c("green","yellow","red"))+
  scale_x_reverse()
```


```{r}
# Shot Index
shot.index <- which(sapply(fr_ar,function(x) x$type$name)=="Shot")
shot.index


# Json Format覺n覺 Direk Data Frame'e D繹n羹t羹rme
shots <- lapply(fr_ar[shot.index], function(x){ unlist(x)})

library(plyr)
shots <- rbind.fill(lapply(shots, 
    function(x) do.call("data.frame", as.list(x))
))
```

```{r}
home_team ="France"
away_team ="Argentina"

shots <- shots %>% 
  filter(type.name == "Shot") %>% 
  select(-contains("freeze")) %>% 
  mutate_if(is.factor, as.character) %>% 
  mutate_at(vars(c(location1, location2, shot.statsbomb_xg)), funs(as.numeric))

sut <- shots %>%  
  mutate(location2 = if_else(team.name == home_team,  80 - location2, location2),
         location1 = if_else(team.name == away_team, 120 - location1, location1)) 


##### XG'羹n neden karesini al覺yor oray覺 tam anlamad覺m eitimde

# Expected Goal
pitch()+
  geom_point(sut %>% filter(shot.outcome.name == "Goal"), mapping = aes(location1, location2, color = team.name, size = sqrt(shot.statsbomb_xg*15)))+
  geom_point(sut %>% filter(shot.outcome.name != "Goal"), mapping = aes(location1, location2, color = team.name,  size = sqrt(shot.statsbomb_xg*15)),  alpha = 0.4)+
  scale_color_manual(values = c("red", "blue"))+
  geom_text(sut %>% filter(shot.outcome.name == "Goal"), mapping = aes(location1, location2, label = player.name),vjust = -0.5, hjust = -0.01)+
  theme(legend.position = "bottom",plot.title = element_text(hjust=0.5, vjust = -5))

```


```{r}
pitch(pitch_theme = "green")+
  geom_point(sut %>% filter(shot.outcome.name == "Goal"), mapping = aes(location1, location2, color = team.name, size = sqrt(shot.statsbomb_xg*15)), shape = 21, fill = "red" )+
  geom_point(sut %>% filter(player.name != "Lionel Andr矇s Messi Cuccittini",shot.outcome.name != "Goal"), mapping = aes(location1, location2, color = team.name,  size = sqrt(shot.statsbomb_xg*15)),  alpha = 0.4)+
  scale_color_manual(values = c("white", "navy"))+
  geom_point(sut %>% filter(player.name == "Lionel Andr矇s Messi Cuccittini"), 
             mapping = aes(location1, location2, size = sqrt(shot.statsbomb_xg*15)))+
  geom_text(sut %>% filter(player.name == "Lionel Andr矇s Messi Cuccittini"), mapping = aes(location1+1, location2-2, label = "L. Messi"),
            size = 2)
```

```{r}
# Pitch
pitch(pitch_theme = "green")+
# Messi's Shots
  geom_point(sut %>% filter(player.name == "Lionel Andr矇s Messi Cuccittini"), 
             mapping = aes(location1, location2, size = sqrt(shot.statsbomb_xg*15)), show.legend = FALSE)+
  geom_text(sut %>% filter(player.name == "Lionel Andr矇s Messi Cuccittini"), mapping = aes(location1+3, location2-2, label = "L. Messi"),
            size = 5)+
# If outcome result is goal
  geom_point(sut %>% filter(shot.outcome.name == "Goal"), mapping = aes(location1, location2, size = sqrt(shot.statsbomb_xg*15)), shape = 21, fill = "red" , show.legend = FALSE)+
# Except Messi's Shots & Goals  
  geom_point(sut %>% filter(player.name != "Lionel Andr矇s Messi Cuccittini",shot.outcome.name != "Goal"), mapping = aes(location1, location2, color = team.name,  size = sqrt(shot.statsbomb_xg*15)),  alpha = 0.4, show.legend = FALSE)+
  scale_color_manual(values = c("white", "navy"))+
# Title
  theme(legend.position = "bottom", text = element_text(color = "white"),plot.title = element_text(hjust=0.5, vjust = -5),
        plot.subtitle = element_text(hjust=0.5, vjust = -5))+
  labs(title = "Argentina | France",
       subtitle = "FIFA World Cup 18")
```







